---
title: "Homework 5"
output: github_document
date: "2025-11-8"
author: Evan Kennedy, ELK2149
---

***Problem 1***


# make initial function, after pulling in necessary libaries for the whole project
```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(broom)
library(tidyverse)
library(janitor)

birthdays = sample(1:365, 5, replace = TRUE)

repeated_bday = length(unique(birthdays)) < 5

repeated_bday
```

#put this in a function

```{r}

bday_sim = function(n_room) {

    birthdays = sample(1:365, n_room, replace = TRUE)

    repeated_bday = length(unique(birthdays)) < n_room

    repeated_bday
    
}

bday_sim(5)
```

#make it from 10000 iterations, group by bdays

```{r}

bday_sim_results = 
  expand_grid(
    bdays = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(
    bdays
    ) |> 
  summarize(
    prob_repeat = mean(result)
  )
  
bday_sim_results
```

#plot this

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line()
```

#the probability of having two people with the same birthday does not increase at a linear pace as the group size increased from 2 to 50 people.

***Problem 2***

```{r}
#set variables
n = 30
sd = 5
u = 0
a = 0.05

#run the single function

run_one = function(u, n, sd) {
  x = rnorm(n, mean = u, sd = sd)
  t_test = t.test(x, mu = 0)
  tibble(
    sample_mean  = mean(x),
    p_value = tidy(t_test)$p.value
  )
}

#run the larger simulation
simulation = 
  expand_grid(iter = 1:5000) |> 
  mutate(out = map(iter, ~ run_one(u = u, n = n, sd = sd))) |>
  unnest(out)

#estimate type 1 error
estimated_type1_error =
  simulation |>
  summarize(type1 = mean(p_value < a)) |>
  pull(type1)

estimated_type1_error

simulation_1 = 
  expand_grid(
    u = 1:6,
    iter = 1:5000
  ) |>
  mutate(out = map(u, ~ run_one(u = .x, n = n, sd = sd))) |>
  unnest(out)

#calculate power
power_results = 
  simulation_1 |>
  mutate(reject = p_value < a) |>
  group_by(u) |>
  summarize(
    power = mean(reject)
  )

power_results
```

```{r}

#graph power results
power_results |>
  ggplot(aes(x = u, y = power)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Value of μ",
    y = "Power"
  )

```

# Power increases as μ increases.

```{r}

#
estimates_1 =
  simulation_1 |>
  mutate(reject = p_value < a) |>
  group_by(u) |>
  summarize(
    mean_hat = mean(sample_mean),
    mean_hat_reject = mean(sample_mean[reject]),
    .groups = "drop"
  )

#double plot of estimate of U and estimate of U only when null was rejected

estimates_1 |>
  ggplot(aes(x = u)) +
  geom_point(aes(y = mean_hat)) +
  geom_line(aes(y = mean_hat)) +
  geom_point(aes(y = mean_hat_reject)) +
  geom_line(aes(y = mean_hat_reject)) +
  labs(
    x = "Value of μ",
    y = "Average estimate of u"
  )

```

#Across all simulations, the average u estimate is approximately the true u. When restricting to rejected tests, the average u is inflated, especially for small u values.

***Problem 3***

```{r}
#first, pull in dataset

homicides = 
  read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") |> 
  clean_names()

#create city_state variable
homicides =
  homicides |>
  mutate(city_state = str_c(city, ", ", state))

homicides |> select(city, state, city_state) |> head()

```

# now collected all unsolved cases for each

```{r}
city_summaries =
  homicides |>
  mutate(unsolved_cases = disposition %in% c("Closed without arrest", "Open/No arrest")) |>
  group_by(city_state) |>
  summarize(
    total = n(),
    unsolved = sum(unsolved_cases),
    .groups = "drop"
  )

city_summaries
```

# focusing on baltimore

```{r}

baltimore =
  city_summaries |>
  filter(city_state == "Baltimore, MD")

baltimore

baltimore_test =
  prop.test(
    x = baltimore$unsolved,
    n = baltimore$total
  )

baltimore_tidy =
  tidy(baltimore_test)

baltimore_tidy |>
  select(estimate, conf.low, conf.high)

```

# now run the function for all cities

```{r}

city_results =
  city_summaries |>
  mutate(
    test = map2(unsolved, total, ~ prop.test(.x, .y))
  ) |>
  mutate(
    test_tidy = map(test, tidy)
  ) |>
  unnest(test_tidy)

city_results =
  city_results |>
  select(city_state, total, unsolved, estimate, conf.low, conf.high)

city_results

```

# now plot this, using error bars

```{r}
city_results |>
  mutate(city_state = fct_reorder(city_state, estimate)) |>
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  coord_flip() +
  labs(
    x = "City",
    y = "Proportion Unsolved",
    title = "Proportion of Unsolved Homicides by City"
  )
```
